---
title: "PA1_template"
author: "Mukesh kumar.P"
date: "March 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```






# **Loading and preprocessing the data**
```{r echo=TRUE}
data<-read.csv("activity.csv")
head(data)
```



# **Total number of steps taken per day**
```{r echo=TRUE}
subset_data=aggregate(data$steps~data$date,FUN=sum)
barplot(subset_data$`data$steps`,names.arg = subset_data$`data$date`,col=subset_data$`data$date`,main="Total number of steps taken per day",xlab="date",ylab="steps")
```




###     Mean of the total number of steps taken per day
```{r}
mean(subset_data$`data$steps`,na.rm = T)
```

###     Median of the total number of steps taken per day
```{r}
median(subset_data$`data$steps`,na.rm = T)
```


# **Average daily activity pattern**
```{r}
subset_data=aggregate(data$steps~data$interval,FUN=mean)
plot(subset_data$`data$interval`,subset_data$`data$steps`,type="l",xlab="Time  interval",ylab="Average steps",col="red",main="Average daily activity pattern")
```
 
###  **5-minute interval containing the maximum number of steps**
```{r}
subset_data$`data$interval`[subset_data$`data$steps`==max(subset_data$`data$steps`)]
```

#  **Inputing missing values**
```{r}
sum(is.na(data$steps))
```

### Strategy for filling in all of the missing values 

    *Fill in the NA values with the mean of the steps in that particular day.*


### New dataset  with the missing data filled.
```{r}
data1<-data  # new dataset
x<-is.na(data1$steps)
subset_data=aggregate(data1$steps~data1$interval,FUN=mean)
names(subset_data)=c("interval","steps")
mynewdata=merge(data1,subset_data,by="interval")
data1$steps[x]=mynewdata$steps.y[x]
sum(is.na(data1$steps))
```
### Total number of steps taken per day
```{r}
subset_data_new=aggregate(data1$steps~data1$date,FUN=sum)
barplot(subset_data_new$`data1$steps`,names.arg=subset_data_new$`data1$date`,col=subset_data_new$`data1$date`,xlab="Date",ylab="Steps",main="Steps taken per day")
```

### NEW MEAN
```{r}
subset_data_new=aggregate(data1$steps~data1$date,FUN=sum)
mean(subset_data_new$`data1$steps`,na.rm=T)
```

### NEW MEDIAN
```{r}
subset_data_new=aggregate(data1$steps~data1$date,FUN=sum)
median(subset_data_new$`data1$steps`,na.rm=T)
```


The impact of the missing data seems rather low, at least when estimating the total number of steps per day.

# **Differences in activity patterns between weekdays and weekends**

```{r}
type_of_day <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        0
    } else {
        1
    }
}
data1$typeday <- as.factor(sapply(data1$date, type_of_day))
head(data1)
```

HERE  1 refers to weekday and 0 refers to weekend


### Average number of steps taken

```{r}
par(mfrow=c(1,2))
newdata<-aggregate(steps~interval,data=data1,subset=(data1$typeday==1),FUN=mean)
plot(newdata$interval,newdata$steps,type="l",xlab="interval",ylab="steps",main="Weekdays")
newdata<-aggregate(steps~interval,data=data1,subset=(data1$typeday==0),FUN=mean)
plot(newdata$interval,newdata$steps,type="l",xlab="interval",ylab="steps",main="Weekends")
```




















